Tropical Andes Taxonomic Diversity of Plants and Frugivores

### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
move some code to L2_3_TropicalAndes_divRelationships

  # Extract cell values
  comparison_data <- data.frame(Plant_Richness = plant_richness_grid$num_species,
                               Frugivore_Richness = frugivore_richness_grid$num_species)

  # Filter out rows where either plant or frugivore richness is zero
  comparison_data_filtered <- comparison_data %>%
                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)

  # Create histograms
  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")
  # Return a list of data
  list(plantgridRichTA = plant_plot,
       frugivoregridRichTA = frugivore_plot,
       plant_cellRichness = plant_richness_grid,
       frugivore_cellRichness = frugivore_richness_grid, 
       comparison_data_filtered = comparison_data_filtered,
       plant_richness_hist = plant_richness_hist,
       frugivore_richness_hist = frugivore_richness_hist)
}

@@

# save plots
plantgridRichTA_1km <- richness_1km$plant_plot
frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
# save richness data
plant_cellRichness_1km <- richness_1km$plant_cellRichness
frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness

comparison_data_filtered_1km <- richness_1km$comparison_data_filtered
```

@@

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness

comparison_data_filtered_5km <- richness_5km$comparison_data_filtered
```

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness

comparison_data_filtered_10km <- richness_10km$comparison_data_filtered
```

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness

comparison_data_filtered_25km <- richness_25km$comparison_data_filtered
```

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness

comparison_data_filtered_50km <- richness_50km$comparison_data_filtered
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness

comparison_data_filtered_75km <- richness_75km$comparison_data_filtered
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness

comparison_data_filtered_100km <- richness_100km$comparison_data_filtered
```

```{r}
# Compare plant and frugivore richness
  ## Fit a linear model
  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
                                                  comparison_data_filtered_100km$Plant_Richness))
  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared

  ## Fit an exponential model
  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
                                                       log(comparison_data_filtered_100km$Plant_Richness)))
  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared

  ## Create scatter plot with linear and exponential lines
  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
  geom_point() +
  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
  geom_smooth(method = "lm", se = FALSE, color = "red")+
  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```

```{r}
# Fit the nonlinear model using nls
nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
                 data = comparison_data_filtered_100km, 
                 start = list(b = 200, c = -0.00001),
                 algorithm = "port",
                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))

# Summary of the model
summary(nls_model)
```

@@

# Compare r-squared by resolution
Plot
```{r}
# create dataframe of resolutions and r-squared values 
#[ADD 1, 5,]
resolution <- c( 10, 25, 50, 75, 100)
#ADD compare_richness_r2_1km, compare_richness_r2_5km,
r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
r_squared_compare <- data.frame(resolution, r_squared)

compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
    geom_point()+
    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
print(compare_r2_by_res)
```

>>>>

  # Create histograms
  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")
  # Return a list of data
  list(plantgridRichTA = plant_plot,
       frugivoregridRichTA = frugivore_plot,
       plant_cellRichness = plant_richness_grid,
       frugivore_cellRichness = frugivore_richness_grid, 
       plant_richness_hist = plant_richness_hist,
       frugivore_richness_hist = frugivore_richness_hist)
}

@@

# save plots
plantgridRichTA_1km <- richness_1km$plant_plot
frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
# save richness data
plant_cellRichness_1km <- richness_1km$plant_cellRichness
frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
```

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
```

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
```

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
```

@@

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
```

@@
